---
title: "Investigating Median Household Income and Potential Influencing Factors in Six States"
author: "Audrey Zhu and Ying Xiong"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: TRUE
---

# Preliminaries

```{r setup, echo=FALSE, cache=FALSE}
## Dr Love recommends you leave this code chunk exactly as it is
library(knitr)
library(rmdformats)

## Global options
options(max.print="100")
opts_chunk$set(comment=NA)
opts_knit$set(width=75)
```

## R Packages

```{r load_packages_here, message = FALSE}
library(janitor)
library(magrittr)
library(naniar)
library(survival)
library(Formula)
library(Hmisc)
library(tidyverse)
```

## Data Ingest

```{r read_in_data_here, message = FALSE}
chr_2021_raw <- read_csv("data/analytic_data2021.csv", skip = 1, guess_max = 4000)

head(chr_2021_raw)
```

# Data Development

## Selecting Data

```{r}
chr_2021 <- chr_2021_raw %>% 
  filter(county_ranked == 1) %>% 
  filter(state %in% c("CA", "MN", "MT", "OH", "WA", "WV")) %>% 
  select(fipscode, state, county, v063_rawvalue, v069_rawvalue, v082_rawvalue,
         v137_rawvalue, v166_rawvalue) %>% 
  rename(median_household_income = v063_rawvalue,
         some_college = v069_rawvalue,
         single_parent = v082_rawvalue,
         long_commute_raw = v137_rawvalue,
         broadband_access_raw = v166_rawvalue) %>% 
  mutate(median_household_income = median_household_income / 1000,
         some_college = some_college * 100,
         single_parent = single_parent * 100,
         long_commute_raw = long_commute_raw * 100,
         broadband_access_raw = broadband_access_raw * 100)

summary(chr_2021$long_commute_raw)
summary(chr_2021$broadband_access_raw)
```

## Repairing the `fipscode` and Factoring the `state`

```{r}
chr_2021 <- chr_2021 %>% 
  mutate(fipscode = str_pad(fipscode, 5, pad = "0"),
         state = factor(state))
```

## Checking the Initial Work

```{r}
glimpse(chr_2021)
```

```{r}
chr_2021 %>% tabyl(state) %>% adorn_pct_formatting()
```

## Creating a binary categorical variable for the percentage of households with a broadband internet connection called broadband_access_2cat

The 2 categories are generated from `broadband_access_raw` and are separated by its median.

```{r, message = FALSE}
chr_2021 <- chr_2021 %>%
    mutate(broadband_access_2cat = case_when(
      broadband_access_raw < median(broadband_access_raw) ~
        "low",
      TRUE ~ "high"),
      broadband_access_2cat = factor(broadband_access_2cat))

mosaic::favstats(broadband_access_raw ~ broadband_access_2cat, data = chr_2021) %>% 
    kable(digits = 3)
```

## Creating a 4-category variable for the percentage of workers who drive to work alone that have a commute longer than 30 minutes called long_commute_5cat

The 5 categories are generated from `long_commute_raw` and are split along its quintiles. 

```{r}

chr_2021 <- chr_2021 %>%
  mutate(long_commute_5cat = case_when(
    long_commute_raw < 20.78 ~ "1_very_low",
    long_commute_raw < 26.92 ~ "2_low",
    long_commute_raw < 33.65 ~ "3_medium",
    long_commute_raw < 41.40 ~ "4_high",
    TRUE ~ "5_very_high"
  ),
  long_commute_5cat = factor(long_commute_5cat))

mosaic::favstats(long_commute_raw ~ long_commute_5cat, data = chr_2021) %>% 
  kable(digits = 3)

names(chr_2021)
nrow(chr_2021)
```

## Structure of the Tibble 

```{r}
str(chr_2021)
```

# Proposal Requirements 

## Proposal Requirement 1 

The 6 states that we selected are California, Minnesota, Montana, Ohio, Washington, and West Virginia. The total number of counties we are studying is 375. The number of counties by state are as follows: 58 for California, 87 for Minnesota, 48 for Montana, 88 for Ohio, 39 for Washington, and 55 for West Virginia. We initially selected Minnesota and Washington because those were the states where the two of us have our permanent residences. We then selected California, Montana, and West Virginia based on the relative level of urbanization in those three states, according to information from [Iowa State University](https://www.icip.iastate.edu/tables/population/urban-pct-states). We noticed that California had the highest percentage of its state population living in an urban area, whereas Montana and West Virginia were among the lowest. 

```{r}
chr_2021 %>% count(state)
```

## Proposal Requirement 2 (Codebook)

Variable | Description
--------- | ------------------------------------------------
fipscode | FIPS code
state | State: the six states are CA, MN, MT, OH, WA, and WV 
county | County name
median_household_income | (v063) Median household income (in thousands), which will be our **outcome** 
some_college | (v069) Percentage of the adult population with some college education
single_parent | (v082) Percentage of children in single-parent households 
long_commute_raw | (v137) Percentage of those who drive to work alone with a commute longer than 30 minutes
broadband_access_raw | (v166) Percentage of households with broadband internet access 
broadband_access_2cat | binary categorical transformation of `broadband_acces_raw`; 2 levels: low = broadband access below 78.99% (i.e. the median), or high
long_commute_5cat | multi-categorical transformation of `long_commute_raw`; 5 levels: 1_very_low = long_commute_raw below 20.78%, 2_low = below 26.92%, 3_medium = below 33.64%, 4_high = below 41.40%, and 5_very high = at or below 69.40% 

We are interested in having the quantitative variable `median_household_income` (`v063`) as our outcome because we believe it is a broad variable that can be influenced by many potential factors and can serve as an intuitive output for our other variables of interest. We selected `some_college` (`v069`), which represents the percentage of the population who has at least some college education, because we hypothesized that it would be correlated with median household income, in which a higher value for `some_college` would predict higher incomes. Our motivation to choose `single_parent` (`v082`), the percentage of children living in a single-parent household, was also due to its potential implications on income; we hypothesize that a higher percentage of single-parent households has a negative correlation with median household income. 

We chose to investigate `broadband_access` (`v166`), the percentage of households with broadband internet access, and our interest in this variable is due to its capacity as a proxy for technological accessibility in a household and its statistical association with income. Broadband access can be potentially related to the outcome in that technological accessibility and high-speed internet at home may enable individuals to find higher-income jobs, while higher incomes would allow people to afford high-speed internet. We turned `broadband_access_raw` from a quantitative variable into a binary categorical variable (`broadband_access_2cat`) based on the cutoff of 78.44%, which is the median percentage, and classified the categories as either `low` or `high`.

We chose to investigate `long_commute` (`v137`), the percentage of people who drive alone to work who have a commute longer than 30 minutes, because we are interested in learning whether commute times can be used as a predictor for median household income. `long_commute` can be related to income in that individuals with long commutes can afford the costs associated with such commutes and thus have higher incomes. Conversely, it is possible that individuals without long commutes, especially those in urban areas, may not be able to afford to live further from their workplace. We turned `long_commute_raw` from a quantitative variable into a multi-categorical variable with 5 categories (`long_commute_5cat`) based on the quintiles, minimum, and maximum of the variable's distribution. The 5 categories are as follows, in ascending order of percentage: the first quintile (0-20%) is `1_very_low`, the second quintile (20-40%) is `2_low`, the third quintile (40-60%) is `3_medium`, the fourth quintile (60-80%) is `4_high`, and the final quintile (80-100%) is `5_very_high`. We chose to divide `long_commute_raw` into 5 categories because this variable's spread is not large enough to justify too many categories, and because we wanted an odd number of categories to avoid splitting the data into a "low" half and a "high" half. We feel that 5 categories appropriately satisfies both these parameters

## Proposal Requirement 3 

```{r}
chr_2021 
```

## Proposal Requirement 4 

```{r}
Hmisc::describe(chr_2021)
```

## Three Important Checks 

```{r}
chr_2021 %>%
  miss_var_summary()

chr_2021 %>% 
  dplyr::summarize(across(median_household_income:broadband_access_raw, ~ n_distinct(.)))

chr_2021 %>% tabyl(broadband_access_2cat)
chr_2021 %>% tabyl(long_commute_5cat)
```

There are no missing values across any of the 10 variables in our tibble `chr_2021`, as observed when we piped `chr_2021` into the `miss_var_summary` function. The raw versions for each of our five selected variables also have at least 10 distinct non-missing values, as seen via the `dplyr::summarize` function. In fact, all of them have hundreds of distinct values for each variable. Finally, for both of our two transformed categorical variables, each level in both factors has at least 10 counties, as observed when we piped `chr_2021` into the `tabyl` function. 

## Saving the Tibble 

```{r}
saveRDS(chr_2021, file = "data/chr_2021_Audrey_Zhu_Ying_Xiong.Rds")
```

## Proposal Requirement 5 

The most challenging part of completing Project A so far was manipulating two of the quantitative variables into categorical variables. Although the code for this process has been already supplied by Dr. Love in the proposal RMarkdown template, we ourselves had to decide what variables made sense to actually transform and what were appropriate cutoffs for both variables. We overcame this difficulty by talking with each other about the variables we were interested in, namely `long_coummute` and `broadband_access`, and by thinking about logical cutoffs for both variables. Our open discussions will be useful in tackling future difficulties in Project A as they arise. 

## Session Information 

```{r}
sessionInfo()
```
