---
title: "Investigating Median Household Income and Potential Influencing Factors in Six States from the 2021 County Health Rankings Dataset"
author: "Audrey Zhu and Ying Xiong"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: TRUE
---

# Preliminaries

```{r setup, echo=FALSE, cache=FALSE}
## Dr Love recommends you leave this code chunk exactly as it is
library(knitr)
library(rmdformats)

## Global options
options(max.print="100")
opts_chunk$set(comment=NA)
opts_knit$set(width=75)
```

## R Packages

```{r load_packages_here, message = FALSE}
library(janitor)
library(magrittr)
library(naniar)
library(survival)
library(Formula)
library(Hmisc)
library(tidyverse)
```

## Data Ingest

```{r read_in_data_here, message = FALSE}
chr_2021_raw <- read_csv("data/analytic_data2021.csv", skip = 1, guess_max = 4000)
head(chr_2021_raw)
```

# Data Development

## Selecting Data

```{r}
chr_2021 <- chr_2021_raw %>% 
  filter(county_ranked == 1) %>% 
  filter(state %in% c("CA", "MN", "MT", "MS", "OH", "WA")) %>% 
  select(fipscode, state, county, v063_rawvalue, v069_rawvalue, v082_rawvalue,
         v023_rawvalue, v166_rawvalue) %>% 
  rename(median_household_income = v063_rawvalue,
         some_college = v069_rawvalue,
         single_parent = v082_rawvalue,
         unemployment = v023_rawvalue,
         broadband_access = v166_rawvalue) %>% 
  mutate(median_household_income = median_household_income / 1000,
         some_college = some_college * 100,
         single_parent = single_parent * 100,
         unemployment = unemployment * 100,
         broadband_access = broadband_access * 100)

head(chr_2021)
summary(chr_2021$unemployment)
summary(chr_2021$broadband_access)
```

## Repairing the `fipscode` and Factoring the `state`

```{r}
chr_2021 <- chr_2021 %>% 
  mutate(fipscode = str_pad(fipscode, 5, pad = "0"),
         state = factor(state))
```

## Checking the Initial Work

```{r}
glimpse(chr_2021)
```

```{r}
chr_2021 %>% tabyl(state) %>% adorn_pct_formatting()
```

## Creating a Binary Categorical Variable for unemployment

[Write Explanation Here] (The median is 4.475)

```{r, message = FALSE}
chr_2021 <- chr_2021 %>%
    mutate(unemployment_2cat = case_when(
      unemployment < median(unemployment, na.rm = TRUE) ~
        "low",
      TRUE ~ "high"),
      unemployment_2cat = factor(unemployment_2cat))

mosaic::favstats(unemployment ~ unemployment_2cat, data = chr_2021) %>% 
    kable(digits = 3)
```

## Creating a 4-Category Variable for broadband_access 

The categories are generated based on information from the minimum, first quartile, median, third quartile, and maximum. 

```{r}
chr_2021 <- chr_2021 %>%
  mutate(broadband_access_4cat = case_when(
    broadband_access < 71.96 ~ "1_very_low",
    broadband_access < 78.44 ~ "2_low",
    broadband_access < 83.07 ~ "3_high",
    TRUE ~ "4_very_high"
  ))

mosaic::favstats(broadband_access ~ broadband_access_4cat, data = chr_2021) %>% 
  kable(digits = 3)

names(chr_2021)
str(chr_2021)
```

# Proposal Requirements 

## Proposal Requirement 1 
The six states that we selected are California, Minnesota, Montana, Mississippi, Ohio, and Washington. The total number of counties we are studying is 402. The number of counties by state are as follows: 58 for California, 87 for Minnesota, 82 for Mississippi, 48 for Montana, 88 for Ohio, and 39 for Washington. We initially selected Minnesota and Washington because those were the states where Ying and Audrey have their permanent residences, respectively. We then selected California, Montana, and Mississippi based on the relative level of urbanization in those three states, according to information from Iowa State University (https://www.icip.iastate.edu/tables/population/urban-pct-states). We noticed California had the highest percentage of its state population living in an urban area while Montana and Mississippi were among the lowest. 

```{r}
chr_2021 %>% count(state)
```

## Proposal Requirement 2 

Variable | Description
--------- | ------------------------------------------------
3 required variables 
fipscode | FIPS code
state | State: my six states are CA, MN, MT, MS, OH, and WA
county | County Name

5 additional variables and transformations for 2 of them 
median_household_income | (v063_rawvalue), quantitative; Median household income in thousands, which will be our **outcome** 
some_college | (v069_rawvalue), quantitative; Percentage of some college education
single_parent | (v082_rawvalue), quantitative; Percentage of children in single-parent households 
unemployment | (v023_rawvalue), later turned to binary categorical; Percentage of unemployment 
broadband_access | (v166_rawvalue), later turned to multi-categorical; Percentage of broadband internet access 
unemployment_2cat | transformed to 2 levels from unemployment: low = unemployment below 4.475% (i.e. the median), or high
broadband_access_4cat | transformed to 4 levels from broadband_access: very low = broadband_access below 71.96%, low = broadband_access below 78.44%, high = broadband_access below 83.07%, and very high = broadband_access at 93.25% or lower. 

We are interested in having the quantitative variable of median household income as our outcome because we believe it is a broad variable that can be influenced by many potential factors and can serve as an intuitive output for our other variables of interest. We selected some college education because we hypothesized that it would be correlated with median household income, in which a higher percentage of some college education predicted higher incomes. Conversely, we were interested in unemployment because we envisioned higher unemployment would negatively impact median household income and vice versa for lower unemployment. Our motivation to choose percentage of children in single-parent households was due to its potential implications on income, in which a higher percentage of single-parent households may negatively affect income. Our last variable was broadband access, and our interest was due to its use as a proxy for technological accessibility in a household and its potential influence on or reflection of income, such as if a family can afford broadband access given their income. Finally, for unemployment, we turned it from a quantitative variable into a binary categorical variable based on the cutoff of 4.475%, which was the median unemployment rate, and classified it as either "low" or "high". For broadband access, we turned it from a quantitative variable into a 4-categorical variable based on 4 groups inspired by the first quartile (71.96%), median (78.44%), third quartile (83.07%), and maximum (93.25%). The 4 categories were as follows: below 71.96% was "very low", below 78.44% was "low", below 83.07% was "high", and at or below 93.25% was "very high". 

## Proposal Requirement 3 

```{r}
chr_2021 
```

## Proposal Requirement 4 

```{r}
Hmisc::describe(chr_2021)
```

## Three Important Checks 

```{r}
chr_2021 %>%
  miss_var_summary()

chr_2021 %>% 
  dplyr::summarize(across(median_household_income:broadband_access, ~ n_distinct(.)))

chr_2021 %>% tabyl(unemployment_2cat)
chr_2021 %>% tabyl(broadband_access_4cat)
```

There are no missing values across any of the 10 variables in our tibble chr_2021. The raw versions for each of our five selected variables also have at least 10 distinct non-missing values. In fact, all of them appear distinct. Finally, for each of our two categorical variables, each level in either of the two factors has at least 10 counties. 

## Saving the Tibble 

```{r}
saveRDS(chr_2021, file = "chr_2021_Audrey_Zhu_Ying_Xiong.Rds")
```

## Proposal Requirement 5 

The most challenging part of completing Project A so far was manipulating 2 quantitative variables into the 2 categorical variables, namely one binary and another multi-category variable. Although the code for this process has been supplied by Dr. Love in the R markdown template, we ourselves had to envision what variables made sense to actually transform and what were appropriate cutoffs for both variables. We overcame this difficulty by talking with each other about variables we were interested in, namely rates of unemployment and broadband access, and thinking about logical cutoffs for both variables. Our open discussions will be useful to tackle future difficulties in Project A as they arise. 

## Session Information 

```{r}
sessionInfo()
```




